Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-10-06 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:10:06	18:49:45	(Tue 6.Oct)

\IEEEPARstart{T}{he} best adaptive sonars are found in nature. Bats, for instance, use a high frequency sonar to detect, identify and track prey with amazing precision in caves amongst a myriad of other bats. Active human made sonars have little to no adaptability on transmit, but several adaptive methods have been investigated post process the received data. Perhaps the most studied method for reconstructing a sonar image is the minimum variance distortion-less response (MVDR) beamformer~\cite{Capon1969}. It adapts the sonar array's response to minimize the influence of noise and interference in the final image.

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This makes MVDR impracticable and sometimes infeasible. We have shown in previous studies that MVDR's complexity can be reduced, as well as accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014,Asen2013}. However, while MVDR relies on estimating and inverting a spatial covariance matrix, there exist a method that avoid this altogether.\todo{samme avsnitt?}

This method is called the Low Complexity Adaptive (LCA) beamformer. It was first proposed by Synnev\aa{}g in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar results to MVDR in a system with focused transmit beam. LCA applies a set of predefined windows and selects the one offering the best noise suppression. It relies on the same optimization criterion as MVDR, and may be considered a version of MVDR with a reduced window solution space. Alternatively, it may be viewed as a delay-and-sum (DAS) beamformer with an adaptive extension. As we will demonstrate, LCA performs surprisingly well because a robust MVDR design tends to create windows that resembles slightly varied versions of conventional ones.

Synnev\aa{}g used a window set comprised of rectangular, Kaiser and inverted Kaiser functions\footnote{May also be called Kaiser-Harris windows.}. Of these one specific Kaiser window was micro-steered. One of the Kaiser windows were micro-steered in 6 different directions. In total he let LCA choose from 12 unique windows, which worked well. But no comparison with MVDR was made, and it was not made clear how LCA performs with more or less windows, and if how steering affects it all.

In this work we apply the LCA method to active sonar imaging. Our overall goal is to provide some intuitive understanding\todo{kanskje ikke helt riktig ord her} of how the window database may be designed. Like Synnev\aa{}g we limit our study to the use of variations of the parametric Kaiser window, since we observed negligible performance difference to using this function or any of the common trigonometric ones. However, we steer all our windows and not just some, and provide reason why the Kaiser window is a suitable one, what variations of Kaiser we should include, and how much it should be steered. Furthermore, unlike Synnev\aa{}g we testing it on data with unfocused transmit beam.

>>>>

\IEEEPARstart{T}{he} best adaptive sonars are found in nature. Bats, for instance, use a high frequency sonar to detect, identify and track prey with amazing precision in caves amongst a myriad of other bats. Active human made sonars have little to no adaptability on transmit, but several adaptive methods have been investigated post process the received data. Perhaps the best studied method for reconstructing a sonar image is the minimum variance distortion-less response (MVDR) beamformer~\cite{Capon1969}. It adapts the sonar array's response to minimize the influence of noise and interference in the final image.

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This makes MVDR impracticable and sometimes infeasible. We have shown in previous studies that MVDR's complexity can be reduced, and its implementation accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014,Asen2013}. However, while MVDR relies on estimating and inverting a spatial covariance matrix, there exist a method that avoid this altogether.\todo{samme avsnitt?}

This method is called the Low Complexity Adaptive (LCA) beamformer. It was first proposed by Synnev\aa{}g in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar results to MVDR in a system with focused transmit beams. LCA applies a set of predefined windows and selects the one offering the best noise suppression. It relies on the same optimization criterion as MVDR, and may be considered a version of MVDR with a reduced window solution space. Alternatively, it may be viewed as a delay-and-sum (DAS) beamformer with an adaptive extension.

Synnev\aa{}g used a window set comprised of rectangular, Kaiser and inverted Kaiser functions\footnote{May also be called Kaiser-Harris windows.}. He made LCA choose from 12 unique windows, 6 of which were micro-steered variants of a fairly wide Kaiser window. While achieving good results, it was not made clear how LCA performs with more or fewer windows, and how steering affects it all.

In this work we apply the LCA method to data from the Kongsberg Maritime HISAS 1030 synthetic aperture sonar (SAS). It has 32 sensors, operates at 100\,kHz and an opening angle of 25$^\circ$. We focus on using LCA with variations of the parametric Kaiser window, but a comparison to using trigonometric windows will also be provided. Unlike Synnev\aa{}g we steer all our windows and not just some, and elaborate on the type of Kaiser windows to include and the method of steering.
